---
title: "DS1_Lab4"
output: html_document
date: "2022-11-09"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(fields)
library(GGally)
library(ggplot2)

```

# Problem 1
```{r}
library(datasets)
data(iris)
summary(iris)
```

```{r}
set.seed(100) 

k_means <- function(data, k, iter){
  # randomly assigning each observation to a cluster
  data <- data %>%
    mutate(cluster = sample(1:k, nrow(data), replace = TRUE))
  # calculate the centroids 
  centroids <- data %>%
    group_by(cluster) %>%
    summarise(across(everything(), mean)) %>%
    select(Sepal.Length:Petal.Width)
  # calculate distance
  points <- data %>%
    select(Sepal.Length:Petal.Width)
  for (i in iter){
    # select the minimum distance
    min_dis<- apply(rdist(centroids, points), 2, which.min)
  }
  return(min_dis) 
}

# Data Wrangling 
data(iris)

iris <- iris %>% select(-Species) 

# Scaling the Data
iris_scaled <- as.data.frame(scale(iris))

k_means(iris_scaled, 3, 50)
```


# Problem 2

```{r message = FALSE, warning=FALSE}
set.seed(2)

# Clustering
cluster <- k_means(iris_scaled, 3, 50)     

# Adding Cluster Value to the Table 
iris_scaled <- mutate(iris_scaled, 
                      cluster_val = cluster)  

iris_scaled$cluster_val <- as.factor(iris_scaled$cluster_val)  

# Plotting Characteristics
ggpairs(iris_scaled, 
        legend = 1, 
        columns = c("Sepal.Length", 
                    "Sepal.Width", 
                    "Petal.Length",
                    "Petal.Width"),
        columnLabels = c("Sepal Length", 
                         "Sepal Width", 
                         "Petal Length", 
                         "Petal Width"),
        aes(color = cluster_val), 
        title="Visualizing Sepal Characteristics by Clusters") + 
  labs(fill='Cluster') 


```

# Problem 3
```{r}

set.seed(100)

wss <- function(k, data) {
  kmeans(data, k, nstart = 50)$tot.withinss
}

k_values <- 1:15
wss_values <- map_dbl(k_values, wss, data = iris_scaled)

wss_values <- data.frame(wss = wss_values,
                         k = k_values)

## Plotting Elbow Curve

g <- ggplot(wss_values, aes(x = k, y = wss)) +
    geom_point() +
  geom_line() + 
  labs(title = "Within-Cluster Sum of Squares vs. Cluster Size \nfor the Iris Data Set",
       y = "Within-Cluster Sum of Squares",
       x = "Cluster Size")
g

```

The appropriate number of clusters for the iris data set is 3 clusters. As evident in the figure above, the "kink" occurs at a cluster size of 3 for the curve hence that being the appropriate cluster size to use. This kink occurs at where the slope of the curve switches from steep to shallow. From this plot, it is clear that additional clusters beyond the third cluster have little value. 


